Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-14 manuscript/manuscript.tex
ms updates

\begin{align}
  \label{eq:conv-corr}
  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
\end{align}
%
where $A$ and $B$ here are two replicate labels, and for the
\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$. Between
replicate populations with identical allele frequency changes, this is one.

@@

Additionally, we wanted to ensure that our temporal covariances and $G(t)$
trajectories were robust to more complicated, but realistic fitness models. To
this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
during burnin, followed by one of two two optima shift routines: (1) sudden
optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and
(2) very graduate optima shifts of $\mu_\text{gradual} = \{0.001,
0.01\}$ per generation using the same two population simulation scheme
described above.  We used a polygenic architecture for these simulations,
\vb{with trait alleles assigned a $\pm 0.01$ effect size with equal
probability, trait mutation rate $10^{-8}$}, and the optima shift began at five
generations after a $10N$ generation burnin. Across our GSS simulations, we
see the expected selection response (Supplementary Material Figure
\ref{suppfig:gss-zbar}).

@@

  \caption{Mean temporal covariance ($\cov \Delta p_5, p_t$, with $t$ varying
    across the x-axis) across 30 replicate simulations (light gray points are
    before the onset of selection; dark gray points are after selection
    begins), under different Gaussian stabilizing selection with optima shift
  regimes. The solid red line is a loess-smoothed average of these points.}

@@

Overall, we see the same qualitative results under Gaussian stabilizing
selection with optima shifts as under exponential directional selection.
Stronger directional selection, here determined by larger sudden optima shifts
or larger gradual shifts per generation, lead to stronger temporal covariances
(Supplementary Materials Figure \ref{suppfig:covs}). Furthermore, we see a
stronger effect of linked selection, as measured by $G(t)$, under stronger
directional selection (Supplementary Material Figure \ref{suppfig:gss-G}).

@@

\vb{Additionally, we looked at the effect the size of each replicate population
  has on a single population's $G(t)$ trajectories. These simulations had the
  same $10N$ generation burnin, followed by a change in population size
  emulating the bottlenecks associated with creating selection lines. Overall,
  we find that smaller population sizes lead to a reduced $G(t)$ (Supplementary
  Material Figure \ref{suppfig:gss-G-sampleN}). This is expected, as the
  denominator of $G(t)$ is $\var(p_t - p_0)$, which has an inverse relationship
  with $N_e$; as replicate population size is reduced, the proportion of allele
  frequency change driven by linked selection is lower, since the rate of drift
  is increased. To isolate the effects of varying replicate population size, we
  also looked at just the magnitude of temporal covariances (Supplementary
  Material Figure \ref{suppfig:gss-covs-sampleN}). We find that smaller
  replicate population sizes lead to \emph{larger} temporal covariances. We
  then looked at the initial trait variance, which as expected, does not vary
  with replicate population size (since all burnin populations had the same
  size). This implies that the linkage disequilibria is higher in smaller
populations, due to founder effects, which has the effect inflating the
temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.}

>>>>

\begin{align}
  \label{eq:conv-corr}
  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
\end{align}
%
where $A$ and $B$ here are two replicate labels, and for the
\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.

@@

Additionally, we wanted to ensure that our temporal covariances and $G(t)$
trajectories were robust to more complicated, but realistic fitness models. To
this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
during burnin, followed by one of two two optima shift routines: (1) sudden
optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and (2) very graduate
optima shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using
the same two population simulation scheme described above (\vb{these shifts are
in standard deviations, since $V_S = 1$}).  We used a polygenic architecture
for these simulations, with trait alleles assigned a $\pm 0.01$ effect size
with equal probability, trait mutation rate $10^{-8}$, and the optima shift
began at five generations after a $10N$ generation burnin.  Across our GSS
simulations, we see the expected selection response (Supplementary Material
Figure \ref{suppfig:gss-zbar}).

@@

  \caption{Mean temporal covariance ($\cov(\Delta p_5, p_t)$, with $t$ varying
    across the x-axis) across 30 replicate simulations (light gray points are
    before the onset of selection; dark gray points are after selection
    begins), under different Gaussian stabilizing selection with optima shift
  regimes. The solid red line is a loess-smoothed average of these points.}

@@

Overall, we see the same qualitative results under Gaussian stabilizing
selection with optima shifts as under exponential directional selection.
Stronger directional selection, here determined by larger sudden optima shifts
or larger gradual shifts per generation, lead to stronger temporal covariances
(Supplementary Materials Figure \ref{suppfig:gss-covs}). Furthermore, we see a
stronger effect of linked selection, as measured by $G(t)$, under stronger
directional selection (Supplementary Material Figure \ref{suppfig:gss-G}).

@@

Additionally, we looked at the effect the size of each replicate population
  has on a single population's $G(t)$ trajectories. These simulations had the
  same $10N$ generation burnin, followed by a change in population size
  emulating the bottlenecks associated with creating selection lines. Overall,
  we find that smaller population sizes lead to a reduced $G(t)$ (Supplementary
  Material Figure \ref{suppfig:gss-G-sampleN}). This is expected, as the
  denominator of $G(t)$ is $\var(p_t - p_0)$, which has an inverse relationship
  with $N_e$; as replicate population size is reduced, the proportion of allele
  frequency change driven by linked selection is lower, since the rate of drift
  is increased. To isolate the effects of varying replicate population size, we
  also looked at just the magnitude of temporal covariances (Supplementary
  Material Figure \ref{suppfig:gss-covs-sampleN}). We find that smaller
  replicate population sizes lead to \emph{larger} temporal covariances. We
  then looked at the initial trait variance, which as expected, does not vary
  with replicate population size (since all burnin populations had the same
  size). This implies that the linkage disequilibria is higher in smaller
populations, due to founder effects, which has the effect inflating the
temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.
